Broom: Safe Region-Based Memory Management for Dataflow Programs

### 2015-11-16 paper/overview2.tex
updated overview

\paragraph{Transferable Regions}
\name's \emph{transferable regions} are an encapsulation of a data-structure
that can be transferred between autonomous entities
(\eg, between two concurrently executing threads or actors).
% Transferable regions are similar to stack regions in that they are memory regions
% that support arbitrary object allocations, but they differ from stack regions in the
% following ways.
Hence, unlike stack regions, transferable regions are not constrained to have a
lexically scoped lifetime.
(Hence, we also refer to them as \emph{dynamic} regions.)
% regions whose lifetimes transcend the boundaries of lexical blocks, functions, or
% even programs.

Furthermore, transferable regions, unlike stack regions, are first class values of \name:
they are considered objects of \C{Region} class, and, like objects of other
classes, they are created using the \C{new} keyword, can be passed as
arguments, stored in data structures, and returned from methods.
As mentioned above, a transferable region is intended to encapsulate a single
data-structure, consisting of a collection of objects with a distinguished root
object of some type \C{T}, which we refer to as the region's \emph{root} object.
The class \C{Region} is parametric over the type \C{T} of this root object.

The \C{Region} constructor expects a \C{Func}
object - a higher-order function argument that returns an object of
root type.  The idea is to execute this function with the newly
created transferable region as its allocation context. The object thus
created inside the transferable region will serve as its root object.
This is illustrated by the following code that creates a transferable region
(\C{r}) of type \C{T}:
\begin{codejava}
  Region<T> r = new Region<T>(() => new T())
\end{codejava}

@@

The protocol for transferable regions is presented as a finite state machine in Fig.~\ref{fig:region-fsm}.
A transferable region starts its lifetime in a \emph{closed} state,
when it is created through \C{new Region} expresion. \name provides

As evident in the above examples, the safety of memory accesses in
\name is now subject to the condition that every transferable region
correctly follows the state transition discipline shown
in Fig.~\ref{fig:region-fsm}. If this is
guaranteed, then \name's region type system statically guarantees the
safety of all memory accesses. In other words, the type system reduces
the problem of ensuring memory safety in \name programs to the problem
of enforcing the state transition discipline for transferable regions.

>>>>

\paragraph{Transferable Regions}
\name's \emph{transferable regions} are an encapsulation of a data-structure
that can be transferred between autonomous entities
(\eg, between two concurrently executing threads or actors).
Hence, unlike stack regions, transferable regions are not constrained to have a
lexically scoped lifetime.
(Hence, we also refer to them as \emph{dynamic} regions.)

Furthermore, transferable regions, unlike stack regions, are first class values of \name:
they are objects of the class \C{Region}, 
they are created using the \C{new} keyword,
and can be passed as arguments, stored in data structures, and returned from methods.
A transferable region is intended to encapsulate a single data-structure,
consisting of a collection of objects with a distinguished root object of some type \C{T},
which we refer to as the region's \emph{root} object.
The class \C{Region} is parametric over the type \C{T} of this root object.

The \C{Region} constructor takes as a parameter a function that constructs
the root object: it creates a new region and invokes this function, with the new
region as the allocation context, to create the root object of the region.
This following code illustrates the creation of a transferable region:
% (\C{r}) of type \C{T}:
\begin{codejava}
  Region<T> r = new Region<T>(() => new T())
\end{codejava}

@@

The protocol for transferable regions is presented as a finite state machine in Fig.~\ref{fig:region-fsm}.
A transferable region starts its lifetime in a \emph{closed} state, when it is created.
%
The safety of memory accesses in \name is now subject to the condition that
every transferable region correctly follows the state transition discipline shown
in Fig.~\ref{fig:region-fsm}. If this is guaranteed, then \name's region type system
statically guarantees the safety of all memory accesses.
% In other words, the type system reduces
% the problem of ensuring memory safety in \name programs to the problem
% of enforcing the state transition discipline for transferable regions.
